---
title: "Select Empty Columns"
author: "Ravi Chowdhury"
date: "12-12-2023"
categories: [tidyverse, code, R]
---

This post is about selecting empty columns using R and `tidyverse`, specifically `dplyr`. There might be a need where you want to select only empty columns from a table for quality check. For example you expect a column to have data and if the column is empty in the table you can perform checks.

With the `janitor` package, using the `remove_empty` function, you can remove empty columns but I am not sure if you can get a list of columns that are empty or have no data.

```{r}
#| label: load packages
#| include: false
library(tidyverse)
```

## Read data

```{r data}
billboard
```

We will create an empty column in this table to reference it and find it.

## Create empty column

```{r}
billboard_modified <- billboard |> 
  mutate(
    empty_col = NA_character_ #empty column
  )
```

## Show / select just the blank or empty columns

```{r}
billboard_modified |> 
  select(where( ~ all(is.na(.))))
```

We will use `colnames()` function to show the list of columns since this table has multiple empty columns.

```{r}
billboard_modified |> 
  select(where( ~ all(is.na(.)))) |> 
  colnames()
```

As you can see it is very easy to find empty columns in a table.
